Prefetching SPARQL query cacher

### 2016-05-26 dissertation/cache-proxy.tex
discuss join rotation

>>>>

While the system used a simple brute-force query planner in the above
evaluation, using an Iterative Dynamic Programming planner
\cite{Kossmann:2000:IDP:352958.352982} would simply amount to
composing \pmodule{Attean::API::IDPJoinPlanner} in the planner class.

While this performed better in most cases, the test suite contains a
query where it does not:
\begin{verbatim}
?s <http://example.org/m/p> "1" ;
   <http://example.org/m/p> ?o ;
   <http://example.org/m/q> _:xyz ;
   <http://example.org/m/q> <http://example.org/m/a> .
?a <http://example.org/m/b> <http://example.org/m/c> .
\end{verbatim}
Note that the above Basic Graph Pattern contains a triple pattern that
does not share a variable (\sparql{?a}) with any other triple pattern,
thus creating a problematic Cartesian join. Additionally, in this
test, the results of the two first triple patterns have been
cached. \textit{A priori}, I have assumed that the best plan is a plan
where the cache is used for these two triple patterns, the last triple
pattern is answered by a Linked Data Fragments server and the last two
triple patterns are coalesced into a new SPARQL query, and evaluated
remotely. This is also coded as a pass criterion for the test. With
the simple query planner, this test passes, but not with the IDP
planner, as it results in a plan where the two triple patterns that
should run is a single SPARQL query is split into two. This resulted
in an interesting observation:

As noted in Section~\ref{sec:implqueryplan}, we rely on rotating joins
to gather triple pattern plans into a larger structure that can be
evaluated together, like a remote SPARQL BGP. On careful inspection, I
found that the join rotation would have had to descend into
great-grandchildren of the top join plan. However, with the IDP
planner, plans are aggressively pruned, and therefore, this
constellation is not considered. This led us (with Gregory Todd
Williams) to conjecture that this weakness is inherit to IDP planning
when using join rotation. We did not attempt to examine this any
further, however.

To remedy this problem, we considered creating a custom query planner
for Basic Graph Patterns, but I did not implement it. The outline
would be as follows:
\begin{enumerate}
\item Find triple patterns that are connected with a common variable,
  and group them to ``components''.
\item If a triple pattern have a cached result, create a plan that
  accesses the cache. 
\item If the triple pattern in the previous step caused a Cartesian
  join to be created within a component, create an equivalent
  additional plan for a SPARQL remote evaluation that does not access
  the local cache.
\item Any remaining components consisting of a single triple pattern
  results in a LDF plans.
\item Any remaining components are turned into SPARQL remote BGP
  plans.
\end{enumerate}

This procedure would allow the cost model to balance the cost between
doing a Cartesian join and doing a remote SPARQL evaluation of data
already in the cache.
